---
title: "Calculation 6 - Mediation Analysis"
author: "Michael Zehetleitner"
date: "2022-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```



```{r}
#stopCluster(cl)
library(bayestestR)
library(tidyverse)
library(brms)
#library(doMC)
#registerDoMC(30)
library(foreach)
library(doParallel)
#registerDoParallel(30)
#library(parallel)
cl <- makeCluster(detectCores() - 1, outfile = "parallelLog.txt")
#cl
#library(doParallel)
registerDoParallel(cl)
```


# BRMS
- mediator = COVCI + BVIS + VCI
## Read in the data
Things to tweak:
- number of iterations: first block
- priors: prior(normal(0, 20), class = b) in three model fit calls
- parallel vs. normal for loop: comment out `forach...` and `i <- as.numeric(...)` and uncomment the for loop in block three

The error - if it occurs - occurs in block 3
```{r}
listOfIterations <- c(5000)
dat <- readRDS( file = "data/clusteredDataWide.rds")
dat <- dat %>%
  filter(DEC_GROUP != "fence-sitter")%>% 
  filter(GENDER_f != "diverse")%>% 
  droplevels()


# Attitude is the sum of VCI, COVCI and BVIS
dat$attitude <- dat$VCI_av + dat$BVIS_av + dat$COVCI_av

dat <- dat %>%
  mutate(across(attitude, ~{(.x - mean(.x)) / sd(.x)}, .names = '{.col}_z'))  # z transform
```


# Model comparisons

## Create the different prediction formulae
```{r}

fileNamePathandNameStub <- "fit objects/mediation/mediation comparison"
fileF <- "full"
fileFHAPL <- "Rossen and best fair-harm-auth-pur-lib"
fileHAPL <- "auth-pur-lib Amin"
fileNull <- "null"

fileNameComponents <- c(fileF, fileFHAPL, fileHAPL, fileNull)

createFileName <- function(path, spec, it, ext){
  paste(path, spec, format(it, scientific = FALSE), ext)
}



# full model 
formulaF.f1 <- bf(attitude_z ~ MFQ_LIBERTY_z + MFQ_HARM_z + MFQ_FAIRNESS_z + MFQ_INGROUP_z + MFQ_AUTHORITY_z + MFQ_PURITY_z + background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f  )
formulaF.f2 <- bf(DEC_GROUP ~ MFQ_LIBERTY_z  + MFQ_HARM_z + MFQ_FAIRNESS_z + MFQ_INGROUP_z + MFQ_AUTHORITY_z + MFQ_PURITY_z + attitude_z + background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f )

# Rossen model
formulaFHAPL.f1 <- bf(attitude_z ~ MFQ_LIBERTY_z + MFQ_HARM_z + MFQ_FAIRNESS_z  + MFQ_AUTHORITY_z + MFQ_PURITY_z + background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f  )
formulaFHAPL.f2 <- bf(DEC_GROUP ~ MFQ_LIBERTY_z  + MFQ_HARM_z + MFQ_FAIRNESS_z  + MFQ_AUTHORITY_z + MFQ_PURITY_z + attitude_z  + background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f )

# Amin model
formulaHAPL.f1 <- bf(attitude_z ~ MFQ_LIBERTY_z    + MFQ_AUTHORITY_z + MFQ_PURITY_z + background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f  )
formulaHAPL.f2 <- bf(DEC_GROUP ~ MFQ_LIBERTY_z     + MFQ_AUTHORITY_z + MFQ_PURITY_z + attitude_z  + background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f )

# Null model
formulaNull.f1 <- bf(attitude_z ~  background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f  )
formulaNull.f2 <- bf(DEC_GROUP ~  attitude_z  + background_politicsLeftRight_z + AGE_f + GENDER_f + EDUCATION_f )


# Formulae in list
formulas.f1 <- list(formulaF.f1, formulaFHAPL.f1, formulaHAPL.f1, formulaNull.f1)
formulas.f2 <- list(formulaF.f2, formulaFHAPL.f2, formulaHAPL.f2, formulaNull.f2)


```
## Fit all model combinations

Fit all model combinations in 2 nested loops
```{r}
for(k in 1:length(listOfIterations)){ # cycle through 1k, 10k, and 100k iterations
  iterations <- listOfIterations[k]
  i <- 0
  #for(i in 1:length(formulas.f1)) {
  foreach (n=1:4, .packages = "brms") %dopar% {
  i <- as.numeric(n)
  
    for(j in 1:length(formulas.f2)){
      
      print(paste("--------------- i: ", i, " j: ", j, " -----------"))
      # Fit the multivariate mediation model
      filename <- createFileName(path = fileNamePathandNameStub, spec =  paste("BOTH att ", fileNameComponents[i], " dec ", fileNameComponents[j]), it = iterations, ext = ".rds")
      print(filename)
      f1 <- formulas.f1[[i]]
      f2 <- formulas.f2[[j]]
      jj <- brm(f1 + f2 + set_rescor(FALSE),
                data = dat, 
                family = c(gaussian, bernoulli),
                prior(normal(0, 20), class = b), # was always 20 but I read maybe priors too loose: https://discourse.mc-stan.org/t/error-in-validate-ll-log-ratios-all-input-values-must-be-finite/27741
                iter = iterations, 
                cores = 4, 
                chains = 4,
                seed = 2324,
                save_pars = save_pars(all = TRUE),
                save_model = paste(filename, "model.txt"),
                silent = TRUE,
                backend = "rstan",
                file = filename)
      print("--- about to add_criterion")
      print(filename)
      jj <- add_criterion(jj, c("loo", "waic"), moment_match = TRUE, reloo = TRUE)
      saveRDS(jj, filename)
      rm(jj)
      
      # Fit the decision component for indirect effects
      filename <- createFileName(path = fileNamePathandNameStub, spec =  paste("DEC ", fileNameComponents[j]), it = iterations, ext = ".rds")
      print(filename)
      jj <- brm(f2,
                data = dat, 
                family = bernoulli,
                prior(normal(0, 20), class = b),
                iter = iterations, 
                cores = 4, 
                chains = 4,
                seed = 224,
                save_pars = save_pars(all = TRUE),
                silent = TRUE,
                backend =  "rstan",
                file = filename)
      print("--- about to add_criterion")
      print(filename)
      jj <- add_criterion(jj, c("loo", "waic"), moment_match = TRUE, reloo = TRUE)
      saveRDS(jj, filename)
      rm(jj)
      
      # Fit the attitude component
      filename <- createFileName(path = fileNamePathandNameStub, spec =  paste("ATT ", fileNameComponents[i]), it = iterations, ext = ".rds")
      print(filename)
      jj <- brm(f1 ,
                data = dat, 
                family = gaussian,
                prior(normal(0, 20), class = b),
                iter = iterations, 
                cores = 4, 
                chains = 4,
                seed = 214,
                save_pars = save_pars(all = TRUE),
                silent = TRUE,
                backend =  "rstan" ,
                file = filename)
      print("--- about to add_criterion")
      print(filename)
      jj <- add_criterion(jj, c("loo", "waic"), moment_match = TRUE, reloo = TRUE)
      saveRDS(jj, filename)
      rm(jj)
      
    }
  }
}
```

## Calculate loo and waic for all 16 model combinations
### Loo the f1+f2 models
```{}
for(k in 1:length(listOfIterations)){ # cycle through 1k, 10k, and 100k iterations
  iterations <- listOfIterations[k]
 # for(i in 1:length(formulas.f1)) {
  i <- 0
  #foreach (n = formulas.f1, .packages = "brms") %do% {
  foreach (n=1:4, .packages = "brms") %dopar% {
    i <- as.numeric(n) 
   #  i <- i+1
    for(j in 1:length(formulas.f2)){
      filename <- createFileName(path = fileNamePathandNameStub, spec =  paste("BOTH att ", fileNameComponents[i], " dec ", fileNameComponents[j]), it = iterations, ext = ".rds")
      print(filename)
      fit <- readRDS(filename)
      try({
        fit <- add_criterion(fit, c("loo", "waic"), moment_match = TRUE, reloo = TRUE)
        saveRDS(fit, filename)
        })
    }
  }
}
```

### Loo the attitude component
```{}
for(k in 1:length(listOfIterations)){ # cycle through 1k, 10k, and 100k iterations
  iterations <- listOfIterations[k]
  for(i in 1:length(formulas.f1)) {
#  i <- 0
 # foreach (n = 1:4, .packages = "brms") %dopar% {
  #  i <- as.numeric(n)
    filename <- createFileName(path = fileNamePathandNameStub, spec =  paste("ATT ", fileNameComponents[i]), it = iterations, ext = ".rds")
    print(filename)
    jj <- readRDS(filename)
    jj <- add_criterion(jj, c("loo", "waic"), moment_match = TRUE, reloo = TRUE)
    saveRDS(jj, filename)
  }
}
```

### Loo the decision component
```{}
for(k in 1:length(listOfIterations)){ # cycle through 1k, 10k, and 100k iterations
  iterations <- listOfIterations[k]
  for(j in 1:length(formulas.f2)){
  # foreach (n = 1:4, .packages = "brms") %dopar% {
   #  j <- as.numeric(n)
    filename <- createFileName(path = fileNamePathandNameStub, spec =  paste("DEC ", fileNameComponents[j]), it = iterations, ext = ".rds")
    print(filename)
    jj <- readRDS(filename)
    jj <- add_criterion(jj, c("loo", "waic"), moment_match = TRUE, reloo = TRUE)
    saveRDS(jj, filename)
    rm(jj)
  }
}
```


# Calculate model weights
Using Bayesian Model Stacking. Takes rather a long computational time.
## For the multivariate f1+f2 fit
```{r}
#listOfIterations <- c(10000)
fileNamePathandNameStub <- "fit objects/mediation/mediation comparison BOTH"
fileF <- "full"
fileFHAPL <- "Rossen and best fair-harm-auth-pur-lib"
fileHAPL <- "auth-pur-lib Amin"
fileNull <- "null"

fileNameComponents <- c(fileF, fileFHAPL, fileHAPL, fileNull)
modelNames <- c(1:16)
 m <- 1
 for(i in 1:4) { # there are 4 tested models for the indirect
    for(j in 1:4){ # and 4 models for the direct effect
      modelNames[m] <- paste("att ", fileNameComponents[i], " dec ", fileNameComponents[j], sep = "")
      m <- m + 1
    }
 }
 modelNames

createFileName <- function(path, spec, it, ext){
  paste(path, spec, it, ext)
}
modelList <- list()
m <- 1


for(k in 1:length(listOfIterations)){ # cycle through 1k, 10k, and 100k iterations
  iterations <- listOfIterations[k]
  for(i in 1:length(formulas.f1)) {
  #foreach (n = 1:4, .packages = "brms") %dopar% {
   # i <- as.numeric(n)
    for(j in 1:length(formulas.f2)){
      filename <- createFileName(path = fileNamePathandNameStub, spec =  paste("att ", fileNameComponents[i], " dec ", fileNameComponents[j]), it = iterations, ext = ".rds")
      print(filename)
      fit <- readRDS(filename)
      modelList[[m]] <- fit
      m <- m + 1
    }
  }
}
list.files("fit objects/with covariates/")
getwd()
modWeight <- model_weights(modelList[[1]], modelList[[2]], modelList[[3]], modelList[[4]], 
                  modelList[[5]], modelList[[6]], modelList[[7]], modelList[[8]], 
                  modelList[[9]], modelList[[10]], modelList[[11]],  modelList[[12]],
                  modelList[[13]], modelList[[14]], modelList[[15]], modelList[[16]],
                  weights = "stacking")
saveRDS(modWeight, file = "fit objects/mediation/modelWeightsBOTH.rds")
```
## Total, direct, and indirect effects

```{r}
# Total effects
total.null <- readRDS(file = "fit objects/with covariates/brmsfit decision cov null 10000 .rds")
total.aut.pur.lib <- readRDS(file = "fit objects/with covariates/brmsfit decision cov auth-pur-lib Amin 10000 .rds")
total.harm.fair.auth.pur.lib <- readRDS(file = "fit objects/with covariates/brmsfit decision cov Rossen and best fair-harm-auth-pur-lib 10000 .rds")
total.full <- readRDS(file = "fit objects/with covariates/brmsfit decision cov full 10000 .rds")


# indirect effects
indirect.null <- readRDS(file = "fit objects/mediation/mediation comparison ATT  null 100000 .rds")
indirect.aut.pur.lib <- readRDS(file = "fit objects/mediation/mediation comparison ATT  auth-pur-lib Amin 100000 .rds")
indirect.harm.fair.auth.pur.lib <- readRDS(file = "fit objects/mediation/mediation comparison ATT  Rossen and best fair-harm-auth-pur-lib 100000 .rds")
indirect.full <- readRDS(file = "fit objects/mediation/mediation comparison ATT  full 100000 .rds")

# direct effects
direct.null <- readRDS(file = "fit objects/mediation/mediation comparison DEC  null 100000 .rds")
direct.aut.pur.lib <- readRDS(file = "fit objects/mediation/mediation comparison DEC  auth-pur-lib Amin 100000 .rds")
direct.harm.fair.auth.pur.lib <- readRDS(file = "fit objects/mediation/mediation comparison DEC  Rossen and best fair-harm-auth-pur-lib 100000 .rds")
direct.full <- readRDS(file = "fit objects/mediation/mediation comparison DEC  full 100000 .rds")

stackingWeight.total <- model_weights(total.null, total.aut.pur.lib, total.harm.fair.auth.pur.lib, total.full, weights = "stacking")
stackingWeight.indirect <- model_weights(indirect.null, indirect.aut.pur.lib, indirect.harm.fair.auth.pur.lib, indirect.full, weights = "stacking")
stackingWeight.direct <- model_weights(direct.null, direct.aut.pur.lib, direct.harm.fair.auth.pur.lib, direct.full, weights = "stacking")

saveRDS(stackingWeight.total, file = "fit objects/mediation/modelWeightsTotal.rds")
saveRDS(stackingWeight.indirect, file = "fit objects/mediation/modelWeightsIndirect.rds")
saveRDS(stackingWeight.direct, file = "fit objects/mediation/modelWeightsDirect.rds")

```

```{r}
stopCluster(cl)
```

